Compositional Certified Resource Bounds\ifshort{\vspace{-1.15cm}}{}

### 2015-04-08 paper/main.tex
drop a few space tricks
reformat some maths

\ifshort{
                    \jan{XXX THIS FIGURE NEEDS TO BE UPDATED XXX}
}{}

@@

The derivation system for the automatic amortized analysis is defined
in~\pref{fig:auto}.  The derivation rules yield judgements of the form
$$
  (\Gamma_B, Q_B); (\Gamma_R, Q_R);
    (\Gamma, Q) \vdash S \dashv (\Gamma', Q').
$$
%
The part $(\Gamma, Q) \! \vdash \! S \! \dashv \! (\Gamma', Q')$ of the judgement
can be seen as a quantitative Hoare triple.  All assertions are split into
two parts, the logical part and the quantitative part.  The
quantitative part $Q$ represents a potential function as a collection
of rational numbers $q_i$ indexed by the index set $I$.  The
logical part $\Gamma$ is left abstract but is enforced by our
derivation system to respect classic Hoare logic constraints.
%
The meaning of this basic judgment is as follows: If $S$ is executed
with starting state $\state$, the assertions in $\Gamma$
hold, $Q' {\ge} 0$, and at least $Q(\state)$ resources are available then the
evaluation does not run out of resources and, if the execution
terminates in state $\state'$, there are at least $Q'(\state')$
resources left and $\Gamma'$ holds for $\state'$.

The judgement is a bit more involved since we have to take into
account the early exit statements \code{break} and \code{return}.
This is similar to classical Hoare triples in the presence of non-linear
control flow. In the judgement, $(\Gamma_B, Q_B)$ is the postcondition
that holds when breaking out of a loop using \code{break}.  % For example, it is set upon entry
% of looping statements by the rule {\sc Q:Loop} and used before
% \code{break} statements by the rule {\sc Q:Break}.
Similarly, $(\Gamma_R, Q_R)$ is the postcondition that holds when
returning from a function call with \code{return}. \iffull{(This becomes clear
in the rules.)}{}

% These judgements correspond to the logic judgements $\Delta; B; R
% \vdash \htriple P S Q$ where each assertion splits in two parts, a
% logical part $\Gamma$ and a quantitative part $Q$.  That means that
% $(\Gamma_B, Q_B)$ is the postcondition of \code{break} statements,
% $(\Gamma_R, Q_R)$ is the postcondition for return statements, and
% $(\Gamma, Q) \vdash S \dashv (\Gamma', Q')$ can be understood as a
% simple Hoare triple.  We leave out the function context $\Delta$ in
% the presentation to avoid cluttering notations.  Instead we assume a
% fixed function context $\Delta$ that is implicit on all judgements.
% Function contexts are treated exactly as in the logic.  % We call
% % precondition the pair $(\Gamma, Q)$ and postcondition the pair
% % $(\Gamma', Q')$.
% The correspondence between judgements of the
% automation and the ones of the quantitative logic is made formal by
% our soundness proof of the automatic amortized analysis at the end of
% this section.

@@

During the analysis, function specifications are quadruples
$(\Gamma_f, Q_f, \Gamma'_f, Q'_f)$ where $\Gamma_f, Q_f$ depend on
$\Vargs$, and $\Gamma'_f, Q'_f$ depend on $\Vret$.  These parameters
are instantiated by appropriate variables on call sites.  A distinctive
feature of our analysis is that it respects the function abstraction:
when deriving a function specification it generates a set of
constraints and the above quadruple; once done, the constraint
set can readily be reused for every call site and the function need
not be analyzed multiple times.
%
Therefore, the derivation rules are parametric in a function context $\Delta$
that we leave implicit in the rules presented here.
% More details can be found in the extended version.

@@

\newcommand{\loopbody}[0]{\ensuremath{\code{if~(x - 10 \geq 0)~\{x=x{-}10; tick(5);\}~else~break;}}}
\begin{figure*}[t]
\vspace{-.5cm}
\centering
\footnotesize
\begin{mathpar}
\inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
{ \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=1cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2.5cm,rightskip=2.5cm, vdots=1cm]
    { \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
      { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
        { \inferrule*[right={\scriptsize (Q:Dec)},leftskip=0cm,rightskip=0cm]
          {
          }
          { (\cdot, B^\text{de}); ((x {\geq} 10,Q^\text{de})  \vdash \code{x = x - 10}
            \dashv (x {\ge} 0, P^\text{de})
          }
        }
        { (\cdot, B^\text{we}); ((x {\geq} 10,Q^\text{we})  \vdash \code{x = x - 10}
          \dashv (\cdot, P^\text{we})
        }
        \and
        \inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
        {
        }
        { (\cdot, B^\text{ti}); (\cdot,Q^\text{ti})  \vdash \code{tick(5)}
          \dashv (\cdot, P^\text{ti})
        }
      }
      { (\cdot, B^\text{sq}); ((x {\geq} 10,Q^\text{sq}) \vdash
        \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{sq})
      }
    }
    { (\cdot, B^\text{if}); ((x {\geq} 10,Q^\text{if}) \vdash
      \code{x=x{-}10; tick(5);} \dashv (\cdot,P^\text{if})
    }
    \and
    \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
    { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
      { }
      { (\cdot, B^\text{br}); (\cdot,Q^\text{br}) \vdash
        \code{break} \dashv (\bot,P^\text{br})
      }
    }
    { (\cdot, B^\text{el}); (x {<} 10,Q^\text{el}) \vdash
      \code{break} \dashv (\cdot,P^\text{el})
    }
  }
  { \hspace{3cm} (\cdot, B^\text{lo}); (\cdot,Q^\text{lo}) \vdash
    \loopbody \dashv (\cdot,P^\text{lo}) \hspace{3cm}
  }
}
{ (\cdot, B); (\cdot, Q) \vdash \code{loop}~{\loopbody} \dashv (\cdot, P)
}
\vspace{-.3cm}
\end{mathpar}
\flushleft
\begin{tabular}{l}Constraints:\end{tabular}
\vspace{-.2cm}
$$
\begin{array}{lll}
P {=} B^\text{lo} \land Q {=} Q^\text{lo} {=} P^\text{lo}
& B^\text{el} {=} B^\text{if} {=} B^\text{lo}
    \land Q^\text{el} {=} Q^\text{if} {=} Q^\text{lo}
    \land P^\text{el} {=} P^\text{if} {=} P^\text{lo}
& B^\text{el} {=} B^\text{br} \land Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}
    \land P^\text{br} \succeq_{(\cdot)} P^\text{el}
\\
B^\text{br} {=} Q^\text{br}
& B^\text{if} {=} B^\text{sq} \land Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}
    \land P^\text{sq} \succeq_{(\cdot)} P^\text{if}
& B^\text{sq} {=} B^\text{we} {=} B^\text{ti} \land Q^\text{sq} {=} Q^\text{we}
    \land P^\text{we} {=} Q^\text{ti} \land P^\text{ti} {=} P^\text{sq}
\\
Q^\text{ti} {=} P^\text{ti} + 5
& B^\text{we} {=} B^\text{de} \land Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}
    \land P^\text{de} \succeq_{(x{\ge}0)} P^\text{we}
& p^\text{de}_{0,10} {=} q^\text{de}_{0,10} + q^\text{de}_{0,x}
    \land p^\text{de}_{0} {=} q^\text{de}_{0}
    \land \forall (\alpha, \beta) \neq (0,10)  . \,
              p^\text{de}_{\alpha,\beta} {=} q^\text{de}_{\alpha,\beta}
\end{array}
$$
\begin{tabular}{l@{\hspace{5em}}l}
  Linear Objective Function: $
    1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
  $
& Constant Objective Function: $
    1{\cdot}q_0 + 11{\cdot}q_{0,10}
  $
\end{tabular}
\vspace{.1cm}
\caption{An example derivation as produced $\toolname$. The constraints are resolved by an off-the-shelf LP solver.}
\label{fig:derivation}
\end{figure*}

@@

In general, the weakening rule can be applied after every syntax directed
rule.  However, it can be left out in practice at some places to
increase the efficiency of the tool.
The weakening operation $\succeq_{\Gamma}$ is defined by the rule
\textsc{Relax}. It is parameterized by a logical context
used to gather information on interval sizes.  For example,
$P^\text{de} \succeq_{(x{\ge}0)} P^\text{we} \equiv$
% SSS SPACING
%\vspace{-.3cm}
{\small
\begin{align*}
  p^\text{we}_{0,10} \leq p^\text{de}_{0,10} + u_{0,10} - v_{0,10}
  \land p^\text{we}_{x,0} \leq p^\text{de}_{x,0} + u_{x,0} - v_{0,x} \land \;\;
  \\
  p^\text{we}_{0} \leq p^\text{de}_{0} - 10 {\cdot} u_{0,10} + 10 {\cdot} v_{0.10}
  \bigwedge_{(\alpha, \beta) \neq (0,10), (x,0)} p^\text{we}_{\alpha,\beta} \leq p^\text{de}_{\alpha,\beta} \;.
\end{align*}
}
% SSS SPACING
%\vspace{-.8cm}

>>>>

\jan{QUENTIN FIX ABOVE FIGURE}

@@

The derivation system for the automatic amortized analysis is defined
in~\pref{fig:auto}.  The derivation rules yield judgements of the form
$$
  (\Gamma_B; Q_B), (\Gamma_R; Q_R) \vdash
    \htriple{ \Gamma; Q }{ S }{ \Gamma'; Q' }.
$$
%
The part $\htriple{ \Gamma; Q }{ S }{ \Gamma'; Q' }$ of the judgement
can be seen as a quantitative Hoare triple.  All assertions are split into
two parts, the logical part and the quantitative part.  The
quantitative part $Q$ represents a potential function as a collection
of rational numbers $q_i$ indexed by the index set $I$.  The
logical part $\Gamma$ is left abstract but is enforced by our
derivation system to respect classic Hoare logic constraints.
%
The meaning of this basic judgment is as follows: If $S$ is executed
with starting state $\state$, the assertions in $\Gamma$
hold, $Q' {\ge} 0$, and at least $Q(\state)$ resources are available then the
evaluation does not run out of resources and, if the execution
terminates in state $\state'$, there are at least $Q'(\state')$
resources left and $\Gamma'$ holds for $\state'$.

The judgement is a bit more involved since we have to take into
account the early exit statements \code{break} and \code{return}.
This is similar to classical Hoare triples in the presence of non-linear
control flow. In the judgement, $(\Gamma_B; Q_B)$ is the postcondition
that holds when breaking out of a loop using \code{break}.  % For example, it is set upon entry
% of looping statements by the rule {\sc Q:Loop} and used before
% \code{break} statements by the rule {\sc Q:Break}.
Similarly, $(\Gamma_R; Q_R)$ is the postcondition that holds when
returning from a function call with \code{return}. \iffull{(This becomes clear
in the rules.)}{}

% These judgements correspond to the logic judgements $\Delta; B; R
% \vdash \htriple P S Q$ where each assertion splits in two parts, a
% logical part $\Gamma$ and a quantitative part $Q$.  That means that
% $(\Gamma_B; Q_B)$ is the postcondition of \code{break} statements,
% $(\Gamma_R; Q_R)$ is the postcondition for return statements, and
% $(\Gamma; Q) \vdash S \dashv (\Gamma'; Q')$ can be understood as a
% simple Hoare triple.  We leave out the function context $\Delta$ in
% the presentation to avoid cluttering notations.  Instead we assume a
% fixed function context $\Delta$ that is implicit on all judgements.
% Function contexts are treated exactly as in the logic.  % We call
% % precondition the pair $(\Gamma, Q)$ and postcondition the pair
% % $(\Gamma', Q')$.
% The correspondence between judgements of the
% automation and the ones of the quantitative logic is made formal by
% our soundness proof of the automatic amortized analysis at the end of
% this section.

@@

During the analysis, function specifications are quadruples
$(\Gamma_f; Q_f, \Gamma'_f; Q'_f)$ where $\Gamma_f; Q_f$ depend on
$\Vargs$, and $\Gamma'_f; Q'_f$ depend on $\Vret$.  These parameters
are instantiated by appropriate variables on call sites.  A distinctive
feature of our analysis is that it respects the function abstraction:
when deriving a function specification it generates a set of
constraints and the above quadruple; once done, the constraint
set can readily be reused for every call site and the function need
not be analyzed multiple times.
%
Therefore, the derivation rules are parametric in a function context $\Delta$
that we leave implicit in the rules presented here.
% More details can be found in the extended version.

@@

\newcommand{\loopbody}[0]{\ensuremath{\code{if~(x \geq 10)~(x=x - 10; tick(5))~else~break}}}
\begin{figure*}[t]
\vspace{-.5cm}
\centering
\footnotesize
\begin{mathpar}
\inferrule*[right={\scriptsize (Q:Loop)},leftskip=0cm,rightskip=0cm,width=20cm]
{ \inferrule*[right={\scriptsize (Q:If)},leftskip=0cm,rightskip=0cm,width=20cm]
  { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=2cm,rightskip=4cm, vdots=1cm]
    { \inferrule*[right={\scriptsize (Q:Seq)},leftskip=0cm,rightskip=0cm]
      { \inferrule*[right={\scriptsize (Q:Weak)},leftskip=0cm,rightskip=0cm]
        { \inferrule*[right={\scriptsize (Q:DecP)},leftskip=0cm,rightskip=0cm]
          {
          }
          { (x {<} 10; B^\text{de}) \vdash
	    \htriple{ x {\geq} 10; Q^\text{de} }{ \code{x = x - 10} }{ \cdot; P^\text{de} }
          }
        }
        { (x {<} 10; B^\text{we}) \vdash
	  \htriple{ x {\geq} 10; Q^\text{we} }{ \code{x = x - 10} }{ \cdot; P^\text{we} }
        }
        \and
        \inferrule*[right={\scriptsize (Q:Tick)},leftskip=0cm,rightskip=0cm]
        {
        }
        { (x {<} 10; B^\text{ti}) \vdash
	  \htriple{ \cdot; Q^\text{ti} }{ \code{tick(5)} }{ \cdot; P^\text{ti} }
        }
      }
      { (x {<} 10; B^\text{sq}) \vdash
	\htriple{ x {\geq} 10; Q^\text{sq} }{ \code{x=x - 10; tick(5)} }{ \cdot; P^\text{sq} }
      }
    }
    { (x {<} 10; B^\text{if}) \vdash
      \htriple{ x {\geq} 10; Q^\text{if} }{ \code{x=x - 10; tick(5)} }{ \cdot; P^\text{if} }
    }
    \and
    \inferrule*[right={\scriptsize (Q:Weak)},leftskip=4.5cm,rightskip=4cm]
    { \inferrule*[right={\scriptsize (Q:Break)},leftskip=0cm,rightskip=0cm]
      { }
      { (x {<} 10; B^\text{br}) \vdash
	\htriple{ x {<} 10; Q^\text{br} }{ \code{break} }{ \bot; P^\text{br} }
      }
    }
    { (x {<} 10; B^\text{el}) \vdash
      \htriple{ x {<} 10; Q^\text{el} }{ \code{break} }{ \cdot; P^\text{el} }
    }
  }
  { \hspace{3cm} (x {<} 10; B^\text{lo}) \vdash
    \htriple{ \cdot; Q^\text{lo} }{ \loopbody }{ \cdot; P^\text{lo} } \hspace{3cm}
  }
}
{ (\cdot; B) \vdash
  \htriple{ \cdot; Q) }{ \code{loop}~{\loopbody} }{ x {<} 10; P }
}
\vspace{-.3cm}
\end{mathpar}
\flushleft
\begin{tabular}{l}Constraints:\end{tabular}
\vspace{-.2cm}
$$
\begin{array}{lll}
P {=} B^\text{lo} \land Q {=} Q^\text{lo} {=} P^\text{lo}
& B^\text{el} {=} B^\text{if} {=} B^\text{lo}
    \land Q^\text{el} {=} Q^\text{if} {=} Q^\text{lo}
    \land P^\text{el} {=} P^\text{if} {=} P^\text{lo}
& B^\text{el} {=} B^\text{br} \land Q^\text{el} \succeq_{(x{<}10)} Q^\text{br}
    \land P^\text{br} \succeq_{(\cdot)} P^\text{el}
\\
B^\text{br} {=} Q^\text{br}
& B^\text{if} {=} B^\text{sq} \land Q^\text{if} \succeq_{(x{<}10)} Q^\text{sq}
    \land P^\text{sq} \succeq_{(\cdot)} P^\text{if}
& B^\text{sq} {=} B^\text{we} {=} B^\text{ti} \land Q^\text{sq} {=} Q^\text{we}
    \land P^\text{we} {=} Q^\text{ti} \land P^\text{ti} {=} P^\text{sq}
\\
Q^\text{ti} {=} P^\text{ti} + 5
& B^\text{we} {=} B^\text{de} \land Q^\text{we} \succeq_{(x{<}10)} Q^\text{de}
    \land P^\text{de} \succeq_{(\cdot)} P^\text{we}
& p^\text{de}_{0,10} {=} q^\text{de}_{0,10} + q^\text{de}_{0,x}
    \land p^\text{de}_{0} {=} q^\text{de}_{0}
    \land \forall (\alpha, \beta) \neq (0,10)  . \,
              p^\text{de}_{\alpha,\beta} {=} q^\text{de}_{\alpha,\beta}
\end{array}
$$
\begin{tabular}{l@{\hspace{5em}}l}
  Linear Objective Function: $
    1{\cdot}q_{x,0} + 10000{\cdot}q_{0,x} + 11{\cdot}q_{x,10} + 9990{\cdot}q_{10,x}
  $
& Constant Objective Function: $
    1{\cdot}q_0 + 11{\cdot}q_{0,10}
  $
\end{tabular}
\vspace{.1cm}
\caption{An example derivation as produced $\toolname$. The constraints are resolved by an off-the-shelf LP solver.}
\label{fig:derivation}
\end{figure*}

@@

In general, the weakening rule can be applied after every syntax directed
rule.  However, it can be left out in practice at some places to
increase the efficiency of the tool.
The weakening operation $\succeq_{\Gamma}$ is defined by the rule
\textsc{Relax}. It is parameterized by a logical context that is
used to gather information on interval sizes.  For example,
{%\small
  \begin{align*}
  P^\text{de} \succeq_{(\cdot)} P^\text{we} &\equiv
 p^\text{we}_{0,10} \leq p^\text{de}_{0,10} + u_{0,10} - v_{0.10} \\
&\land  p^\text{we}_{0} \leq p^\text{de}_{0} - 10 {\cdot} u_{0,10} + 10 {\cdot} v_{0.10} \\
&\land  \forall (\alpha, \beta) \neq (0,10)  . \, p^\text{we}_{\alpha,\beta} \leq p^\text{de}_{\alpha,\beta} \;.
  \end{align*}
}
